---
title: "Olist db"
subtitle: "Monetization Analysis Graded Task"
author: "Leonardo Cerliani"
date: "8/23/2023"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    highlight: pygments
    theme: cerulean
    css: styles.css
---


```{r, message=F}

library(tidyverse)
library(bigrquery)
library(wesanderson)
library(plotly)
library(hrbrthemes)
library(patchwork)
library(flextable)
library(ggstatsplot)
library(GGally)
library(viridis)
library(lubridate)
library(geosphere)

project_id <- "tc-da-1"
dataset_id <- "olist_db"

con <- dbConnect(
  bigrquery::bigquery(),
  project = project_id,
  dataset = dataset_id
)

# to close the connection: dbDisconnect(con)

```


```{sql connection=con, output.var="df_freight_ratio"}

select
  price, freight_value,
  round(freight_value / (freight_value + price),2) as freight_ratio,
from
  `olist_db.olist_order_items_dataset`

```


# Distribution of freight ratio

```{r, warning=FALSE}

df <- df_freight_ratio

# # Simple histogram - refined below by calculating the values manually
# hist(df$freight_ratio, main = "Number of order in different ranges of shipping cost",
#           xlab = "freight ratio = freight cost / (price + freight cost)",
#           ylab = "number of orders", col = "steelblue", border = "white")
# abline(v = 0.19, col = "red")
# text(x = 0.19, y = par("usr")[4] - 2500, labels = "Median = 0.19", pos = 4, col = "red")


# Calculate revenue and number of orders - as well as their proportion and 
# cumulative sum - for different segments of freight ratio
fr_bin_value <- 0.05

ddf <- df %>%
  mutate(fr_bin = cut(freight_ratio, breaks = seq(0, 1, by = fr_bin_value)) ) %>%
  group_by(fr_bin) %>% 
  mutate(fr_bin = max(freight_ratio) ) %>% 
  reframe(
    revenue = sum(price),
    n_orders = n(),
    prop_orders = n() / nrow(df),
    prop_total_revenue = sum(price) / sum(df$price)
  ) %>%
  na.omit() %>% 
  arrange(fr_bin) %>% 
  mutate(
    cum_revenue = cumsum(revenue),
    cum_prop_revenue = cumsum(prop_total_revenue)
  )



scale_dual_axis = 0.2

ddf %>% 
  ggplot(aes(x = fr_bin)) +
  geom_bar(aes(y = prop_orders), stat = "identity", fill = "steelblue", color = "white", width = fr_bin_value) +
  geom_line(aes(y = cum_prop_revenue*scale_dual_axis), color = "red") +
  scale_y_continuous(
    sec.axis = sec_axis(~./scale_dual_axis, name = "% Cumulative Revenue", breaks = seq(0, 1, 0.1))
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(0, 1, fr_bin_value)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(xlim = c(0, 0.8)) +
  labs(
    title = "Orders for different segments of Freight Ratio", 
    x = "Freight ratio segments",
    y = "Proportion of total Orders"
  ) + theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )


```





```{r}
df %>% 
 mutate(price_qtile = ntile(price, 4)) %>% 
 group_by(price_qtile) %>% 
 mutate(price_range = paste0(min(price),"-",max(price)) %>% factor) %>% 
ggplot(aes(x = freight_ratio, fill = as.factor(price_range)) ) +
  geom_histogram(binwidth = 0.02, alpha = 1) +
  # stat_bin(aes(y=..count../sum(..count..)),geom="step", binwidth = 0.02) +
  # geom_density(alpha = 0.7) +
  # facet_wrap(~ price_range, ncol = 2) +
  labs(title = "Histogram of Freight Ratio by Price Range",
       x = "Freight Ratio", 
       y = "Number of Orders",
       fill = "Segments of orders \nby value (quartiles)") +
  theme_modern() +
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")) +
  theme(legend.position = c(0.7,0.7))



# p <- df %>% 
#  mutate(price_qtile = ntile(price, 4)) %>% 
#  group_by(price_qtile) %>% 
#  mutate(price_range = paste0(min(price),"-",max(price)) %>% factor) %>% 
# ggplot(aes(x = freight_ratio, fill = as.factor(price_range)) ) +
#   # geom_histogram(binwidth = 0.02, alpha = 1) +
#   # stat_bin(aes(y=..count../sum(..count..)),geom="step", binwidth = 0.02) +
#   geom_density(alpha = 0.7) +
#   # facet_wrap(~ price_range, ncol = 2) +
#   # labs(title = "Histogram of Freight Ratio by Price Range",
#   #      x = "Freight Ratio", 
#   #      y = "Number of Orders",
#   #      fill = "Segments of orders \nby value (quartiles)") +
#   theme_void() +
#   scale_fill_manual(values = wesanderson::wes_palette("Zissou1")) +
#   theme(legend.position = c(0.7,0.7))
# 
# p + guides(fill = "none")
# 
# ggsave("/Users/leonardo/Dropbox/turing_college/Endorsement/plot.pdf", plot = p, device = "pdf")  # Adjust width and height as needed

```


```{r}

# Freight ratio in 4 categories of products of increasing price
df %>% 
 mutate(price_qtile = ntile(price, 4)) %>% 
 group_by(price_qtile) %>% 
 mutate(price_range = paste0(min(price),"-",max(price)) %>% factor) %>% 
 ggplot(aes(x = price_range, y = freight_ratio)) +
 geom_boxplot(color = "steelblue") +
 stat_summary(
    fun.data = function(y) data.frame(y = median(y), label = median(y)),
    geom = "text", hjust = 4, vjust = 0.5, color = "black", size = 3
  ) +
 theme_minimal() +
 labs(
   title = "Freight ratio for orders of different value",
   x = "Price range (quartiles)",
   y = "Freight ratio" 
 )
```



```{r}

# Freight ratio decrease very quick with order value
df %>% 
 ggplot(aes(x = price, y = freight_ratio)) +
 geom_point(alpha = 0.1, color = "steelblue") +
 # theme_minimal() +
 theme_minimal() +
 scale_x_continuous(
  trans = "log", 
  labels = function(x) round(x,2),
  breaks = 2^(0:12)
 ) + 
  labs(
    title = "Very high shipping fees for orders < 100 BR$",
    x = "Order value (log)",
    y = "Freight ratio"
  )


# # Alternative representation: 2D histogram (to highlight counts)
# df %>% 
#   # filter(freight_ratio >= 0.2) %>%
#   ggplot(aes(x = freight_ratio, y = price)) +
#   geom_hex(bins = 30) +
#   scale_fill_continuous(type = "viridis", trans = "log", breaks = 10^seq(4)) +
#   scale_y_continuous(trans = "log", breaks = 10^seq(3) ) +
#   scale_x_continuous(breaks = seq(0,1,0.1)) +
#   theme_minimal() +
#   coord_fixed(ratio = 0.1) +
#   theme(
#     axis.text.x = element_text(size = 12),
#     axis.text.y = element_text(size = 12),
#     panel.grid.minor = element_blank()
#   )


# # Original (not log transformed) values
# df %>%
#  filter(freight_ratio > 0, price < 1500) %>%
#  sample_frac(0.1) %>%
#  ggplot(aes(x = price, y = freight_ratio)) +
#  geom_point(alpha = 0.1, color = "steelblue") +
#  theme_minimal()

```


```{r, include=F, eval=F}

# Orders count and % of revenue for different value segments

price_binwidth = 10 # in BR$

ddf <- df %>%
  filter(price <= 300) %>%
  mutate(price_bin = cut(price, breaks = seq(0, max(price), by = price_binwidth)) ) %>%
  group_by(price_bin) %>% 
  reframe(
    bin_price = max(price),
    order_count = n(),
    revenue = sum(price),
    prop_orders = n() / nrow(df),
    prop_total_revenue = sum(price) / sum(df$price)
  ) %>%
  select(-price_bin) %>% 
  arrange(bin_price) %>% 
  mutate(
    cum_revenue = cumsum(revenue),
    cum_prop_revenue = cumsum(prop_total_revenue)
  )


scale_dual_axis = 10000

ddf %>% 
  ggplot(aes(x = bin_price)) +
    geom_bar(aes(y = order_count), stat = "identity",
             fill = "steelblue", color = "white") +
    geom_line(aes(y = cum_prop_revenue*scale_dual_axis), color = "red") +
    theme_minimal() +
    scale_y_continuous(
      name = "Order Count",
      sec.axis = sec_axis(~./scale_dual_axis, name = "% Cumulative Revenue", breaks = seq(0,1,0.1))) +
    scale_x_continuous(breaks = c(seq(10,100,10),seq(120,180,20),seq(200,500,50)) ) +
    labs(
      title = "Order count and % of total revenue in different value segments",
      x = "Price bin"
    )

# # Orders < 100 BR$ make for 27% of the revenue
# df$price[df$price < 100] %>% sum() / df$price %>% sum 
# 
# # and 64% of all orders
# df$price[df$price < 100] %>% length / df$price %>% length 

```








# Effect of lowering the freight ratio

## Preliminary check
```{r, include=F, eval=F}
# Preliminary check: 
# price values are log-normally distributed at every freight-ratio segment

df %>%
  filter(freight_ratio > 0) %>% 
  # mutate(fr_05 = round(freight_ratio*20)/20) %>%
  group_by(freight_ratio) %>% 
  ggplot(aes(x = freight_ratio, y = log(price), group = freight_ratio)) + 
  geom_boxplot(color = "steelblue") + 
  theme_minimal()

# df %>%
#   group_by(freight_ratio) %>%
#   reframe(
#     AOV_log = (sum(log(price))/n()) %>% exp,
#     AOV = sum(price)/n(),
#     AOV_median = median(price),
#     AOV_mean = mean(price),
#     cnt = n()
#   )

```



## Revenues after lowering FR of a given percent
```{r}

pct_decrease_fr <- 0.05

ddf <- df %>% 
  rename(fr = freight_ratio) %>%
  filter(fr > 0) %>%
  mutate(lower_fr = round(fr * (1 - pct_decrease_fr),2) ) %>% 
  add_count(fr, name = "cnt_fr") %>% 
  add_count(lower_fr, name = "cnt_lower_fr") %>% 
  mutate(order_increase = cnt_lower_fr / cnt_fr) %>% 
  group_by(fr) %>% 
  reframe(
    AOV = sum(price) / n(),
    current_revenue = round(cnt_fr * AOV,2) %>% unique,
    new_revenue = round(cnt_lower_fr * AOV,2) %>% unique
  ) %>%
  # express in deciles instead of percentiles of freight_ratio
  mutate(ntile_fr = ntile(fr,10)) %>%
  group_by(ntile_fr) %>% 
  mutate(fr_range = paste(min(fr)*100,"-", max(fr)*100,"%")) %>% 
  group_by(fr_range) %>% 
  reframe(
    current_revenue = sum(current_revenue),
    new_revenue = sum(new_revenue)
  ) 

pct_increase_revenue <- round((sum(ddf$new_revenue) / sum(ddf$current_revenue) - 1) * 100,2)

ddf %>% 
  pivot_longer(cols = c("current_revenue","new_revenue"), names_to = "revenue") %>% 
  ggplot(aes(x = fr_range, y = value, fill = revenue)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
        axis.title.y = element_text(angle = 90)) +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(
    title = paste0("Expected increase in revenue: ",pct_increase_revenue,"%"),
    subtitle = paste0("when freight ratio is lowered by ",pct_decrease_fr*100,"%"),
    x = "Freight ratio segment", y = "revenue per segment"
  ) +
  theme(
    legend.position = c(0.7,0.7),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )


```

## Profits after lowering FR

Olist generally takes a cut of 10% on the price of the item (shipping fees excluded).


```{r}

pct_decrease_fr <- 0.05

ddf <- df %>% 
  rename(fr = freight_ratio) %>%
  filter(fr > 0) %>%
  mutate(freight_discount = freight_value * pct_decrease_fr) %>% 
  mutate(lower_fr = round(fr * (1 - pct_decrease_fr),2) ) %>% 
  add_count(fr, name = "cnt_fr") %>% 
  add_count(lower_fr, name = "cnt_lower_fr") %>% 
  mutate(order_increase = cnt_lower_fr / cnt_fr) %>% 
  group_by(fr) %>% 
  reframe(
    AOV = sum(price) / n(),
    AO_freight_discount = sum(freight_discount) / n(),
    current_revenue = round(cnt_fr * AOV,2) %>% unique,
    new_revenue = round(cnt_lower_fr * AOV,2) %>% unique,
    freight_loss = round(cnt_lower_fr * AO_freight_discount,2) %>% unique
  ) %>%
  # express in deciles instead of percentiles of freight_ratio
  mutate(ntile_fr = ntile(fr,10)) %>%
  group_by(ntile_fr) %>% 
  mutate(fr_range = paste(min(fr)*100,"-", max(fr)*100,"%")) %>% 
  group_by(fr_range) %>% 
  reframe(
    current_revenue = sum(current_revenue),
    new_revenue = sum(new_revenue),
    total_freight_loss = sum(freight_loss)
  )


# Applying the freight discount on all freight ratio segments
ddf %>% 
  mutate(
    current_profit = current_revenue * 0.10,
    new_profit = (new_revenue * 0.10) - total_freight_loss
  ) %>% 
  summarise(
    pct_increase_profit = (sum(new_profit) - sum(current_profit)) / sum(current_profit) * 100,
    profit_increase = sum(current_profit) * pct_increase_profit / 100
  ) %>% flextable() %>% 
  add_header_row(values = c("Apply discount to all segments",""))


# Applying the freight discount only to the 10-18% freight ratio
ddf %>% 
  mutate(
    current_profit = current_revenue * 0.10,
    new_profit = ifelse(
      fr_range == "10 - 18 %", new_revenue * 0.10, current_revenue * 0.10 
    )
  ) %>% 
  summarise(
    pct_increase_profit = (sum(new_profit) - sum(current_profit)) / sum(current_profit) * 100,
    profit_increase = sum(current_profit) * pct_increase_profit / 100
  ) %>% flextable() %>% 
  add_header_row(values = c("Apply discount to 10-18% fr segment",""))



```


# Freight ratio and stars

There is actually no difference in the reviews for sellers with different freight ratio

```{sql connection=con, output.var="df_FR_stars"}

select
  t1.order_id,
  round(t1.freight_value / (t1.freight_value + t1.price),2) as freight_ratio,
  review_score as n_stars
from `olist_db.olist_order_items_dataset` t1 join `olist_db.olist_order_reviews_dataset` t2
  on t1.order_id = t2.order_id

```


```{r, warning=FALSE, message=F}

# df_FR_stars %>% 
#   mutate(qtile_FR = ntile(freight_ratio, 4)) %>%
#   group_by(qtile_FR, n_stars) %>% 
#   reframe(
#     cnt = n()
#   ) %>% 
#   ungroup %>% group_by(qtile_FR) %>%
#   reframe(
#     n_stars = unique(n_stars),
#     prop = cnt / sum(cnt)
#   ) 

ddf_stars <- df_FR_stars %>% 
  mutate(qtile_FR = ntile(freight_ratio, 4)) %>% 
  add_count(qtile_FR, name = "n_orders") %>% 
  group_by(qtile_FR, n_stars) %>% 
  reframe(
    prop = unique(n() / n_orders)
  )


# # Quartiles of FR 
# ddf_stars <- df_FR_stars %>% 
#   mutate(qtile_FR = ntile(freight_ratio, 4))
# 
# ggbarstats(
#   y = qtile_FR,
#   x = n_stars,
#   data = ddf_stars,
#   results.subtitle = F,
#   label = "percent",
#   caption = F
# )

# High and low FR
df_FR_stars %>% 
  mutate(FR_segment = ifelse(freight_ratio <= 0.2, "low","high")) %>% 
  ggbarstats(
    y = FR_segment,
    x = n_stars,
    results.subtitle = F,
    label = "percent"
  ) +
  # scale_fill_viridis_d()
  scale_fill_brewer(palette = "Dark1", direction = -1) + # Set1,2,3, Dark1,2,3, Paired, Accent, Pastel1,2, 
  theme(axis.text.x = element_text(size = 12)) +
  labs(
    title = "Number of Stars in Reviews for high and low freight ratio",
    x = "Freight ratio"
  )

```






# Distance between customer and seller

## Get the data using SQL
```{sql connection=con, output.var="df_sql"}
/* Distance between seller and customer */

-- NB the lat/lng for each zipcode are averaged

with 
geo as
(
  select
    distinct(geolocation_zip_code_prefix) zipcode,
    avg(geolocation_lat) as lat,
    avg(geolocation_lng) as lng,
  from `olist_db.olist_geolocation_dataset`
  group by geolocation_zip_code_prefix
  order by
    geolocation_zip_code_prefix
),

seller_info as 
(
  select
    distinct(sellers.seller_id),
    seller_zip_code_prefix as seller_zipcode,
    geo.lat as seller_lat, 
    geo.lng as seller_lng,
    seller_city,
    seller_state,
    order_id,
    price,
    freight_value,
    product_weight_g,
    product_length_cm * product_height_cm * product_width_cm as product_volume_cm3
  from `olist_db.olist_sellers_dataset` sellers join `olist_db.olist_order_items_dataset` items
    on sellers.seller_id = items.seller_id
  join geo on sellers.seller_zip_code_prefix = geo.zipcode
  join `olist_db.olist_products_dataset` products on items.product_id = products.product_id
),
customer_info as
(
  select
    distinct(cust.customer_id),
    customer_zip_code_prefix as customer_zipcode,
    customer_city,
    customer_state,
    order_id,
    date(order_purchase_timestamp) as date_purchased,
    date(order_delivered_customer_date) as date_delivered,
    geo.lat as customer_lat,
    geo.lng as customer_lng
  from 
    `olist_db.olist_customesr_dataset` cust join `olist_db.olist_orders_dataset` orders
      on cust.customer_id = orders.customer_id
    join geo on cust.customer_zip_code_prefix = geo.zipcode
)
select 
  *
from seller_info join customer_info
  on seller_info.order_id = customer_info.order_id
;

```



## Calculate distance in km between seller and customer
```{r}

library(lubridate)

df <- df_sql %>% 
  select(-order_id_1) %>%
  mutate(fr = freight_value / (price + freight_value)) %>%  # calculate freigth_ratio 
  mutate(fr_segment = ifelse(fr <= 0.2, "low","high")) %>%  # set high and low fr (<>0.2)
  mutate(days_to_deliver = interval(date_purchased, date_delivered) %>% as.numeric()  / (60 * 60 * 24)) %>% 
  filter(!is.na(days_to_deliver))

# df %>% glimpse

library(geosphere)

# Function to calculate distance using distGeo
# s_lng/lat = seller; c_lng/lat = customer
calculate_distance <- function(s_lng, s_lat, c_lng, c_lat) {
  distm(
    matrix(c(s_lng, s_lat), ncol = 2), 
    matrix(c(c_lng, c_lat), ncol = 2), 
    fun = distGeo
  )
}

D_vector <- list(
  df$seller_lng, df$seller_lat, 
  df$customer_lng, df$customer_lat
) %>% 
  pmap_dbl(calculate_distance, .progress = TRUE)

df$distance_km <- round(D_vector/1000)

df$distance_km[df$distance_km < 1000] %>% hist


```


## Limit distance to 2000 km and days to deliver to 40
```{r}

df %>% select_if(is.numeric) %>%  summary

describe <- function(column, take_log = 0) {
  par(mfrow = c(1,2))
  var <- df[[column]]
  if(take_log == 1) {var = log(var)}
  boxplot(var, main = column)
  hist(var)
}


describe("days_to_deliver", 0)
describe("fr", 0)

df

df %>% glimpse()

df <- df %>%
  filter(days_to_deliver <= 40) %>%  # filter all orders that took more than 2 months to be delivered
  filter(distance_km < 2000)

# df$days_to_deliver %>% hist(main = "")
# 
# df %>%
#   select(price) %>%
#   mutate(var = price) %>%
#   ggplot(aes)

```


## Association between distance and delivery time
With relation to freight ratio


```{r}
plot(distance_km ~ days_to_deliver, data = df %>% sample_frac(0.01))

boxplot(df$days_to_deliver)
boxplot(df$distance_km)

fit <- lm(days_to_deliver ~ poly(distance_km,2), data = df)
summary(fit)


# # Regular scatterplot. The correlation is invisible
# df %>% 
#   sample_frac(0.5) %>% 
#   select(distance_km, days_to_deliver) %>% 
# ggplot(aes(x = distance_km, y = days_to_deliver)) +
#   geom_point(color = "steelblue", alpha = 0.7) +
#   theme_minimal()


# 2D histogram. The correlation becomes evident
# See
# https://stackoverflow.com/questions/54092169/how-to-plot-ggplots-hexagon-only-if-number-greater-than-threshold
df %>% 
  sample_frac(1) %>% 
  # filter(fr_segment == "high") %>% 
  select(distance_km, days_to_deliver, fr_segment) %>% 
  mutate(distance_bin = cut(distance_km, breaks = seq(0, 2500, by = 10))) %>% 
  group_by(distance_bin, fr_segment) %>%
  mutate(distance_bin = max(distance_km)) %>%
ggplot(aes(x = distance_bin, y = days_to_deliver, 
           fill = cut(..count.., 2^seq(4,15)) )) +
  geom_hex(bins = 30) +
  scale_fill_viridis_d(labels = list(2^seq(4,11),2^seq(5,12)) %>% pmap(~ paste0(.x, "-", .y))) +
  # scale_fill_continuous(type = "viridis", trans = "log", breaks = 2^seq(15)) +
  theme_modern() +
  coord_fixed(ratio = 2000 / 40) +
  labs(
    title = "Number of orders delivered in N days for each distance",
    subtitle = "by freight ratio segment",
    x = "Distance (km)",
    y = "Days to deliver",
    fill = "Number of\nOrders"
  ) +
  facet_wrap(vars(fr_segment), ncol = 2) +
  theme(strip.text = element_text(size = 12))

```


```{r}
# # Continous version - difficult to threshold
# df %>% 
#   sample_frac(1) %>% 
#   select(distance_km, days_to_deliver) %>% 
#   mutate(distance_bin = cut(distance_km, breaks = seq(0, 2500, by = 10))) %>% 
#   group_by(distance_bin) %>%
#   mutate(distance_bin = max(distance_km)) %>%
# ggplot(aes(x = distance_bin, y = days_to_deliver )) +
#   geom_hex(bins = 30) +
#   scale_fill_continuous(type = "viridis", trans = "log", breaks = 2^seq(15)) +
#   theme_minimal() +
#   coord_fixed(ratio = 50) +
#   labs(
#     title = "Number of orders delivered in N days for each distance",
#     x = "Distance (km)",
#     y = "Days to deliver",
#     fill = "Number of\norders"
#   )

```


## Freight ratio and days to delivery
```{r, warning=FALSE, message=FALSE}
df %>%
  sample_frac(1) %>% 
  select(fr, days_to_deliver, fr_segment) %>% 
  mutate(delivery_times_pctile = ntile(days_to_deliver, 5)) %>% 
  ggbarstats(
   y = fr_segment,
   x = delivery_times_pctile,
   label = "percent",
   results.subtitle = F
  ) +
  scale_fill_brewer(palette = "Dark1", direction = 1, labels = c("Very Slow","Slow","Average","Fast","Very Fast")) +
  theme(axis.text.x = element_text(size = 12)) +
  labs(
    title = "Delivery times for high and low freight ratio",
    x = "Freight ratio segment" 
  ) +
    guides(fill = guide_legend(title = "Delivery time"))  # Change "New Legend Title" to your desired title
```


## Differences in price
Smaller orders have much higher freight ratio, but that is not surprising, and we already knew it.

```{r, message=F, warning=F}

df %>% 
  sample_frac(1) %>% 
  select(price, fr, fr_segment) %>% 
  filter(fr > 0) %>%
  ggbetweenstats(
    x = fr_segment,
    y = price,
    results.subtitle = F
  ) +
  scale_y_continuous(trans = "log") +
  theme(axis.text.x = element_text(size = 12))

```




## Linear modelling freight ratio - Exploration

```{r, message=F}

d_sample <- df %>% sample_frac(0.05)

# Explore using only on a sample of df_distance to save time
d_sample %>% 
  filter(fr > 0) %>% 
  select(fr, distance_km, product_weight_g, product_volume_cm3, days_to_deliver) %>% 
  na.omit() %>% 
  ggpairs()


d_sample_std <- d_sample %>% 
  mutate_at(c("fr", "distance_km", "product_weight_g", "product_volume_cm3", "days_to_deliver"), .funs=scale)

df2fit <- df %>% 
  select(fr, distance_km, product_weight_g, product_volume_cm3, days_to_deliver, price) %>%
  filter(fr > 0) %>% 
  mutate(log_fr = log(fr))



# Modelling freight_value also gives a good result
fit <- lm(freight_value ~ distance_km + 
                          product_weight_g +
                          product_volume_cm3 +
                          # days_to_deliver + 
                          # fr +
                          log(price), 
                          data = df) 
fit %>% summary
car::vif(fit)



# Modelling freight_ratio
# Price is by far the most important factor (negatively correlated!): 19% of the variance

fit <- lm(fr ~ price, data = df2fit)
summary(fit)


# We get close to 80% of explained variance!.
fit <- lm(fr ~ 
            distance_km +
            product_weight_g +
            # product_volume_cm3 +
            # days_to_deliver +
            log(price),
            data = df2fit) 

fit %>% summary
car::vif(fit)
# plot(fit)




```



## Linear modelling freight ratio - Dig deeper
```{r}
# df2fit <- df %>% 
#   select(fr, distance_km, product_weight_g, product_volume_cm3, days_to_deliver, price) %>%
#   filter(fr > 0) %>% 
#   mutate(log_fr = log(fr))


# We get > 80% of explained variance!.
fit <- lm(fr ~ 
            log(distance_km + 1) +
            product_weight_g +
            product_volume_cm3 +
            # log(days_to_deliver + 1) +
            log(price),
            data = df2fit) 

fit %>% summary
car::vif(fit)
# plot(fit)
```




```{r}
# Let's make a df with the log of the values we are interested and explore
# the correlations

df2fit <- df %>%
  filter(fr > 0, distance_km > 0, price > 0, price < 2000) %>% 
  mutate(log_fr = log(fr), log_distance_km = log(distance_km), log_price = log(price), fr_pct = fr*100) %>% 
  select(fr, fr_pct, freight_value, price, distance_km, log_fr, log_price, log_distance_km, product_weight_g, product_volume_cm3) %>% 
  filter(log_price > 0, log_distance_km > 0)
  
fit <- lm(
  fr ~ 
    log_price +
    distance_km +
    product_weight_g +
    product_volume_cm3,
  data = df2fit
)  

# log_price explains ~ 64% of the variance in fr
fit <- lm(fr ~ log_price, data = df2fit)

# distance_km explains about 3.5% of the variance
fit <- lm(fr ~ distance_km, data = df2fit)

# together they explain ~ 72% of the variance
fit <- lm(fr ~ log_price + I(distance_km^(0.5)), data = df2fit)

# adding weight and volume explains an additional 5%, to 77% of the variance
# however this leads to a certain extent of violation of lm assumptions

# the best model is obtained with log(1/price) and the sqrt(distance_km) 78% of the variance
fit <- lm(fr ~ log(I(1/log_price)) + I(distance_km^(0.5)), data = df2fit)

# however the most important evidence is that fr is predicted almost entirely from price
# wrt other expected variables, such as distance, weight and volume, so we will use the 
# simplest model:
fit <- lm(fr ~ log_price, data = df2fit)

fit %>% summary
# car::vif(fit)
# plot(fit)


```


Now freight_value
```{r, message=F}
df2fit <- df %>%
  filter(fr > 0, freight_value > 0, distance_km > 0, price > 0, price < 2000) %>%
  mutate(log_fr = log(fr), log_freight_value = log(freight_value),
         log_distance_km = log(distance_km), log_price = log(price)) %>%
  select(fr, freight_value, log_freight_value, price, distance_km, 
         log_fr, log_price, log_distance_km, product_weight_g, product_volume_cm3) %>%
  filter(log_price > 0, log_distance_km > 0)
  

# Modelling freight_value also gives a good result
fit <- lm(freight_value ~ distance_km + 
                          product_weight_g +
                          product_volume_cm3 +
                          log(price), 
                          data = df2fit) 

fit %>% summary
car::vif(fit)

# note also the positive association between price and dimensions 
lm(price ~ product_volume_cm3 + product_weight_g, data = df2fit) %>% summary

fit %>% summary
# car::vif(fit)
# plot(fit)


# freight ratio ggpairs
df2fit %>% 
  sample_frac(0.01) %>% 
  select(log_fr, log_price, distance_km, product_volume_cm3, product_weight_g) %>% 
  ggpairs() +
  theme_minimal() +
  ggtitle("Freight Ratio associations")


# freight value ggpairs
df2fit %>% 
  sample_frac(0.01) %>% 
  select(log_freight_value, log_price, distance_km, product_volume_cm3, product_weight_g) %>% 
  ggpairs() +
  theme_minimal() +
  ggtitle("Freight Value associations")


lm(fr ~ poly(distance_km,2), data = df2fit) %>% summary

```




```{r}

df2fit %>% 
  summarise(
    mean_fr = mean(fr),
    median_fr = median(fr),
    mean_fv = mean(freight_value),
    median_fv = median(freight_value)
  )


df2fit %>% 
  filter(price <= 100) %>% 
  arrange(desc(freight_value)) %>% 
  select(price, freight_value)
  ggplot(aes(x=freight_value)) +
  geom_histogram(bins = 50) +
  theme_minimal()

```




```{r}

df2fit %>% 
  sample_frac(0.01) %>% 
  ggplot(aes(x=fr)) + 
  geom_histogram(bins=30, color = "white", fill = "steelblue") +
  theme_modern()

df2fit %>% 
  select(fr, price, log_distance_km) %>% 
  mutate(log_distance_km_ntile = ntile(log_distance_km + 1, 4)) %>% 
  sample_frac(0.01) %>% 
  filter(fr > 0) %>% 
  ggplot(aes(x=fr, fill = log_distance_km_ntile %>% factor)) + 
  geom_histogram(bins=30, color = "white" ) +
  theme_modern() 

```


































# Garbage collector

```{r, include=F, eval=F}

df %>% 
  select(-freight_value) %>% 
  rename(fr = freight_ratio) %>%
  filter(fr > 0) %>%
  mutate(lower_fr = round(fr * 0.95,2) ) %>% 
  add_count(fr, name = "cnt_fr") %>% 
  add_count(lower_fr, name = "cnt_lower_fr") %>% 
  mutate(order_increase = cnt_lower_fr / cnt_fr) %>%
  group_by(fr) %>% 
  mutate(AOV = sum(price)/n()) %>%
  ungroup() %>% 
  mutate(new_profit = round(cnt_lower_fr * AOV,2)) %>% 
  reframe(
    current_fr = unique(fr),
    AOV = unique(AOV),
    new_profit = unique(new_profit)
  ) %>% arrange(current_fr)




# df %>%
#   group_by(freight_ratio) %>% 
#   mutate(cnt_orders = n()) %>% 
#   mutate(AOV = sum(price) / n()) %>% 
#   mutate(total_profit = AOV * cnt_orders) %>% 
#   reframe(
#     tt = sum(price),
#     total_profit = unique(total_profit)
#   ) %>% reframe(
#     total_profit = sum(total_profit)
#   )
```





```{r, include=F, eval=F}

# I try to see what would happen for a range of fr_discounts
# When applied to all fr segments, it quickly goes negative.
# When applied only to the segment of 10-18% it increases enormously,
# because I assume 100% conversion across all segments. This is not plausible
# though, and to fix it would require too much time, so for the moment I wll
# just leave it as an experiment

estimate_profit <- function(df, pct_decrease_fr) {
  df %>% 
    rename(fr = freight_ratio) %>%
    filter(fr > 0) %>%
    mutate(freight_discount = freight_value * pct_decrease_fr) %>% 
    mutate(lower_fr = round(fr * (1 - pct_decrease_fr),2) ) %>% 
    add_count(fr, name = "cnt_fr") %>% 
    add_count(lower_fr, name = "cnt_lower_fr") %>% 
    mutate(order_increase = cnt_lower_fr / cnt_fr) %>% 
    group_by(fr) %>% 
    reframe(
      AOV = sum(price) / n(),
      AO_freight_discount = sum(freight_discount) / n(),
      current_revenue = round(cnt_fr * AOV,2) %>% unique,
      new_revenue = round(cnt_lower_fr * AOV,2) %>% unique,
      freight_loss = round(cnt_lower_fr * AO_freight_discount,2) %>% unique
    ) %>%
    # express in deciles instead of percentiles of freight_ratio
    mutate(ntile_fr = ntile(fr,10)) %>%
    group_by(ntile_fr) %>% 
    mutate(fr_range = paste(min(fr)*100,"-", max(fr)*100,"%")) %>% 
    group_by(fr_range) %>% 
    reframe(
      current_revenue = sum(current_revenue),
      new_revenue = sum(new_revenue),
      total_freight_loss = sum(freight_loss)
    ) %>% 
    mutate(
      current_profit = current_revenue * 0.10,
      new_profit = ifelse(fr_range == "10 - 18 %", new_revenue * 0.10, current_revenue * 0.10), # only 10-18
      # new_profit = (new_revenue * 0.10) - total_freight_loss  # all FR segments
    ) %>% 
    summarise(
      pct_decrease_fr,
      pct_increase_profit = (sum(new_profit) - sum(current_profit)) / sum(current_profit) * 100,
      profit_increase = sum(current_profit) * pct_increase_profit / 100
    )
}


max_fr_discount <- 0.3

profits <- seq(0, max_fr_discount, 0.01) %>% map_dfr(~ estimate_profit(df,.x))


profits %>% 
  ggplot(aes(x = pct_decrease_fr, y = pct_increase_profit)) +
  geom_line() +
  theme_minimal()

```







## Other non-numeric variables
Suggestion: try doing the heatmaps using reactable
```{r, include=F, eval=F}

df %>% select_if(is.character) %>% summary

ddf <- df %>%
  filter(fr > 0) %>% 
  select(seller_state, customer_state, fr, distance_km) %>% 
  group_by(seller_state, customer_state) %>% 
  reframe(avg_fr = median(fr), avg_D = median(distance_km)) %>% 
  arrange(seller_state, customer_state)


p1 <- ddf %>% 
ggplot(aes(x = customer_state, y = seller_state, fill = avg_fr)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  theme_minimal() +
  labs(
    title = "Heatmap of Average Freight Ratio",
    x = "Customer State",
    y = "Seller State",
    fill = "Average Freight Ratio"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_minimal() +
  coord_fixed(ratio = 1)


p2 <- ddf %>% 
ggplot(aes(x = customer_state, y = seller_state, fill = avg_D)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  theme_minimal() +
  labs(
    title = "Heatmap of Average Distance",
    x = "Customer State",
    y = "Seller State",
    fill = "Average Distance"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_minimal() +
  coord_fixed(ratio = 1)

p1 + p2


```

















